---
title: "Strings and Factors"
output: github_document
---

```{r setup}
library(tidyverse)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
```

## String Vectors

Strings are case sensitive. 
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "m")
str_replace(string_vec, "jeff", "Jeff")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```
`^` indicates beginning of the line -- looking for this pattern at the beginning only
`$` looks at end of line of string



```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"bush")
str_detect(string_vec,"[Bb]ush")

```
`[Bb]` gets upper and lower case bush, can also include a range of values


```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```


```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```
This is looking for 7 followed by anything and then 11.
`.` looks for any character


```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")

```
Use double backslash `\\` for special characters like periods or brackets 


## Why factors are weird

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

as.numeric(factor_vec)

factor_vec = fct_relevel(factor_vec, "male")
as.numeric(factor_vec)
```

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>%
  slice(-1)
```

Let's clean this up

```{r}
marj_df = table_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Midwest", "Northeast", "South", "West")))
```

Do dataframe stuff

```{r}
marj_df %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Restaurant Inspections

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  pivot_wider(names_from = grade, values_from = n)

rest_inspec %>% 
  janitor::tabyl(boro, grade)
```

```{r}
rest_inspec =
  rest_inspec %>% 
  filter(
    str_detect(grade, "[ABC]"),
    !(boro == "Missing")
  ) %>% 
  mutate(boro = str_to_title(boro))
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Iz][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro)
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

What about changing factor labels...

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Iz][Zz][Zz][Aa]")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")
  ) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```
`str_replace` changed factor back to string
